---
title: "LMR_PGPT"
author: "Pramod Verma"
date: "03 May 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
Loading all the required packages

```{r message=FALSE, warning=FALSE}

#install.packages(c("ggplot2", "reshape2","lattice", "corrplot"))
library(ggplot2) ## Grammar of graphics
#library(reshape2) ## Reshaping data frames
#library(lattice) ## More graphics
library(corrplot) ## For plotting the corelations

```

## Reading the file from the computer, change the path accordingly
## GDPE is a transformed variable: GDP + Export

```{r }

file ="D:/Projects/Global Container Throughput Index/Data/Final_Pred.csv"
ModelData = read.csv(
  file = file, 
  header = TRUE,
  colClasses = c("factor", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric"))

# To remove the decimal places from the year values
ModelData$Period <- as.factor(as.integer(as.character(ModelData$Period)))

# End of the file reading into the memory
```
## Check the dataset if everything is correct

```{r }

# Checking the structure of the input data
str(ModelData)

# Checking the first five rows
head(ModelData, 5)

# Summary of the input data
summary(ModelData)

# Checking the corelation
m <- cor(na.omit(ModelData[,-1]))
corrplot(m, method = "number", type="upper")

```

## Defining the training and forecast data

```{r }

# training data
mt <- ModelData[1:18, -1]

# forecast data of predictors
mf <- ModelData[19, -1]

# Training Data
mt

#Forecast Data for independent variables
mf

```


## Building Models with different combination of variables: All Variables

```{r }

# Selecting all the independent variables
fit_all <- lm(PGPT ~ ., data = mt)

# Keeping fit_all in a common variable for repeatability of code
fit <- fit_all

#Checking the summary
summary(fit)
drop1(fit, test = "F")

#Plotting the residuals
qplot(.fitted, .stdresid, data = fit, geom = "hex") + geom_smooth(size = 1) +
xlab("fitted values") + ylab("standardized residuals")

plot(fit)

# Predict the PGPT values using the model
pred <- predict(fit, newdata=mf, interval="confidence", level=.95)
pred1<-pred



```

## Building Models with different combination of variables: Only corelated Variables GDP+Export+Stock+AirLog+CPI

```{r }

# Selecting corelated variables
fit_cor <- lm(PGPT ~ GDPE+Stock+AirLog+CPI, data = mt)

# Keeping fit_all in a common variable for repeatability of code
fit <- fit_cor

#Checking the summary
summary(fit)
drop1(fit, test = "F")

#Plotting the residuals
qplot(.fitted, .stdresid, data = fit, geom = "hex") + geom_smooth(size = 1) +
xlab("fitted values") + ylab("standardized residuals")

plot(fit)

# Predict the PGPT values using the model
pred <- predict(fit, newdata=mf, interval="confidence", level=.95)
pred2 <- pred


```


## Building Models with Export and GDP

```{r }

# Selecting corelated variables
fit_3 <- lm(PGPT ~ GDPE, data = mt)

# Keeping fit_all in a common variable for repeatability of code
fit <- fit_3

#Checking the summary
summary(fit)
drop1(fit, test = "F")

#Plotting the residuals
qplot(.fitted, .stdresid, data = fit, geom = "hex") + geom_smooth(size = 1) +
xlab("fitted values") + ylab("standardized residuals")

plot(fit)

# Predict the PGPT values using the model
pred <- predict(fit, newdata=mf, interval="confidence", level=.95)
pred3 <- pred

```

## Building Models with Export

```{r }

# Selecting corelated variables
fit_4 <- lm(PGPT ~ Export, data = mt)

# Keeping fit_all in a common variable for repeatability of code
fit <- fit_4

#Checking the summary
summary(fit)
drop1(fit, test = "F")

#Plotting the residuals
qplot(.fitted, .stdresid, data = fit, geom = "hex") + geom_smooth(size = 1) +
xlab("fitted values") + ylab("standardized residuals")

plot(fit)

# Predict the PGPT values using the model
pred <- predict(fit, newdata=mf, interval="confidence", level=.95)
pred4 <- pred
rbind(pred1, pred2, pred3, pred4)

```

## Building Models with Transformed Variables : GDP + Export and all others

```{r }

# Selecting corelated variables
fit_5 <- lm(PGPT ~ .- GDP - Export, data = mt)

# Keeping fit_all in a common variable for repeatability of code
fit <- fit_5

#Checking the summary
summary(fit)
drop1(fit, test = "F")

#Plotting the residuals
qplot(.fitted, .stdresid, data = fit, geom = "hex") + geom_smooth(size = 1) +
xlab("fitted values") + ylab("standardized residuals")

plot(fit)

# Predict the PGPT values using the model
pred <- predict(fit, newdata=mf, interval="confidence", level=.95)
pred5 <- pred
rbind(pred1, pred2, pred3, pred4, pred5)


# Predicting the historical values of throughput based on the model and comparing it with the actual historical data

comp <- cbind(as.data.frame(ModelData$Period)[1:18,], mt$PGPT, as.data.frame(fitted(fit)))
comp
```


