---
title: "R Programming Introduction to Workshop Participants"
author: "Pramod Verma, Data For Excellence (DFE), GPE, PSA International"
date: "24 September 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

## <span style="color:blue">**Use of this document**</span> 

This file is created for the sole use of PSA Data Analytics Technical Workshop participants for demonstration and learning about the R programming language. All rights reserved.  

## <span style="color:blue">**Scripting language used**</span> 

This document is created using R Markdown, a scripting language available as open source from R Foundation. 

Loading all the required packages
```{r message=FALSE, warning=FALSE}
#install.packages("readxl",repos = "http://cran.us.r-project.org")
library(readxl)


#install.packages("dplyr",repos = "http://cran.us.r-project.org")
library(dplyr)

```

May need to load more libraries/packages depending on local computer/server
       

## <span style="color:blue">**Loading the file into R data-frame**</span> 

```{r }
#Reading the csv file
port = read.csv('D:/Data Analytics Workshop/Data Analytics Technical Workshop Singapore/Data/port.csv')

#Reading the excel file
#port_excel = read_excel('D:/Data Analytics Workshop/Data Analytics Technical Workshop Singapore/Data/port.xlsx')

```
      

## <span style="color:blue">**Analysing the dataset**</span> 

```{r }
#Looding at the dimension of the dataset (number of rows and columns)
dim(port)

#Looding at the class of the dataset
class(port)

#Looding at the names of the column of the dataset
names(port)


#Looding at the first few rows
head(port)

#Looking at all the columns and their types
str(port)

#Looding at the summary of the entire dataset
summary(port)

```
     
## <span style="color:blue">**Creating Variables in R**</span> 

```{r }
#Simple Variables
x = 5
y = 'Singapore Terminal'

x
y

# List

a = c(2,3,4,5,8,9)
b = c('Singapore', 'Korea', 'India')
c = rnorm(20)


a
b
c
```
    
## <span style="color:blue">**Basic Data Preparation Activities in R **</span> 

### <span style="color:green">**Selecting Variables **</span> 

```{r }
# Take the selected columns in new dataset
newdata = select(port, CNTR_N, COMMODITY, CARGO_OWNER)
head(newdata)

# keep the container number and all variables between logistics partner and cargo owner
newdata = select(port, CNTR_N, LOGISTICS_PARTNER:CARGO_OWNER)
head(newdata)

# keep the container number and all variables except reefer temperature and batch id
newdata = select(port, -BATCH_ID, -RF_TEMP)
head(newdata)


```
  
### <span style="color:green">**Selecting Observations **</span> 

```{r }
# filter for the selected record in the dataset
newdata = filter(port, CNTR_N=='TEXU 1046448')
head(newdata)

newdata = filter(port, CARGO_OWNER=='Airbus')
head(newdata)

# Multiple type of operators can be used to filter records
newdata = filter(port, is.na(RF_TEMP))
head(newdata)

newdata = filter(port, WT_Q > 20000)
head(newdata)

```
  
### <span style="color:green">**Creating/Recoding variables **</span> 

```{r }
# changes the weight to ton
newdata = mutate(port, WT_Q = WT_Q/1000)
head(newdata)

# create a new variable to store weight in ton
newdata = mutate(port, WT_Q_TON = WT_Q/1000)
head(newdata)

# create a new flag to indicate heavy containers
newdata = mutate(port, HEAVY_FLAG = ifelse(WT_Q > 20000, 
                                     "Heavy", 
                                     "Not Heavy"))
head(newdata)

```
  
### <span style="color:green">**Summarizing Data **</span> 

```{r }
# Calculate the mean length and weight of all the containers
newdata = summarise(port, MEAN_LENGTH = mean(LENGTH_Q, na.rm=TRUE), 
                     MEAN_WEIGHT = mean(WT_Q, na.rm=TRUE))
head(newdata)


# Calculate the mean length and weight of all the containers by commodity
newdata = group_by(port, COMMODITY)
newdata = summarise(newdata, MEAN_LENGTH = mean(LENGTH_Q, na.rm=TRUE), 
                     MEAN_WEIGHT = mean(WT_Q, na.rm=TRUE))
newdata

```
    
### <span style="color:green">**Missing Data **</span> 

```{r }
# what is the proportion of missing data for each variable
Percent_Miss <- colSums(is.na(port))/nrow(port)
as.data.frame(Percent_Miss)

```
 

## End of the Script