---
title: "R Visualization Capabilities Introduction"
author: "Pramod Verma, Data For Excellence (DFE), GPE, PSA International"
date: "05 March 2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

## <span style="color:blue">**Use of this document**</span> 

This file is created for the sole use of PSA Data Analytics Technical Workshop participants for demonstration and learning about the R visualization capabilities. All rights reserved.

## <span style="color:blue">**Scripting language used**</span> 

This document is created using R Markdown, a scripting language available as open source from R Foundation. 

Loading all the required packages
```{r message=FALSE, warning=FALSE}

install.packages("dplyr",repos = "http://cran.us.r-project.org")
library(dplyr)

install.packages("ggcorrplot",repos = "http://cran.us.r-project.org") 
library(ggcorrplot)

install.packages("ggplot2",repos = "http://cran.us.r-project.org") 
library(ggplot2)

```

May need to load more libraries/packages depending on local computer/server
       

## <span style="color:blue">**Loading the file into R data-frame**</span> 

```{r }
#Reading the csv file
port = read.csv('D:/Data Analytics/Data Analytics Workshop/Data Analytics Technical Workshop/Version 2/Data/port.csv')


```
      

  
## <span style="color:blue">**Visualization using ggplot library **</span> 

### <span style="color:green">**Scatter Plot **</span> 

```{r }

plot(head(port$DWELL_DAYS,500))

```
    
### <span style="color:green">**Plotting of categorical and continouous variable **</span> 

```{r }
# specify dataset and mapping
table(port$COMMODITY)

#Display as data frame
as.data.frame(table(port$COMMODITY))

# Arrange the data frame
arrange(as.data.frame(table(port$COMMODITY)),desc(Freq))

#plot the dataframe
df = arrange(as.data.frame(table(port$COMMODITY)),desc(Freq))
names(df) = c('COMMODITY', 'FREQ')
g = ggplot(df, aes(x=COMMODITY, y=FREQ))
g+ geom_col() +coord_flip()

#Another simple way of plotting
plot(table(port$COMMODITY))

```
  
### <span style="color:green">**Visualization of co-relation among variables**</span> 

```{r }

#Plot the corelation
ggcorrplot(cor(port[,c('DWELL_DAYS', 'WT_Q')]))
ggcorrplot(cor(port[,c('DWELL_DAYS', 'WT_Q', 'LENGTH_Q')]))
ggcorrplot(cor(port[,c('DWELL_DAYS', 'WT_Q', 'LENGTH_Q')]),lab=TRUE)

```
  
### <span style="color:green">**Looking at the histogram of dwelldays and weight**</span> 

```{r }

#plotting the histogram
hist(port$DWELL_DAYS, col='green')
hist(port$WT_Q, col='red')

#plotting the histogram for a cargo owner
hist(port[port$CARGO_OWNER == 'Renault',]$DWELL_DAYS, col='green')
hist(port[port$CARGO_OWNER == 'Renault',]$WT_Q, col='red')

```
  
### <span style="color:green">**Looking at the boxplot of dwelldays and weight**</span> 

```{r }

#plotting the histogram
boxplot(port$DWELL_DAYS, col='green')
boxplot(port$WT_Q, col='red')

#plotting the histogram for a cargo owner
boxplot(port[port$CARGO_OWNER == 'Dell',]$DWELL_DAYS, col='green')
boxplot(port[port$CARGO_OWNER == 'Dell',]$WT_Q, col='red')

```
  

## <span style="color:red">**End of the Script**</span> 